Oregon_Wildfire_reduced

### 2017-02-21 r_scripts/oregon_clean.rmd
Update oregon_clean.rmd

## The overall summary  of data set
72 variables in total.
```{r overall summary}

# Explore data structure
summary(oregon_df)
```

## 1) *personkey*, Unique person identifier
person's ID, maximum is 14381640
numeric; length: 8
```{r personkey}
summary(as.factor(oregon_df$personkey))
min(oregon_df$personkey)
max(oregon_df$personkey) # seems length is 8
```

>>>>

## The overall summary  of data set
72 variables in total.
```{r overall summary}
# Check data structure
# summary(oregon_df)
```

## convert variable type to appropriate functional form 
(i.e. zipcode leave as character, age as numeric, date as date, dx leave as character, but *NULL* == NA)
```{r}
oregon_df$yob <- as.numeric(oregon_df$yob)
oregon_df[oregon_df=="*NULL*"] <- NA
```

# summary statistics

## 1. Personkey ?? not shown?
```{r}
personkey2 <- oregon_df %>% 
  select(personkey) %>%
  mutate(personkey_cat = ifelse(!is.na(personkey), "Non-Missing", 
                         ifelse(is.na(personkey), "Missing", NA))) %>% 
  group_by(personkey_cat) %>%
  summarise(n = n()) %>%
  mutate(total_n =sum(n),
         percent = round(n/total_n, 4)*100) %>%
      select(personkey_cat, n, percent)

# table ----
knitr::kable(personkey2, caption = "Patient ID")

## 2. Patid ???
```{r}
patid2 <- oregon_df %>% 
  select(patid) %>%
  mutate(patid_cat = ifelse(!is.na(patid), "Non-Missing", 
                         ifelse(is.na(patid), "Missing", NA))) %>% 
  group_by(patid_cat) %>%
  summarise(n = n()) %>%
  mutate(total_n =sum(n),
         percent = round(n/total_n, 4)*100) %>%
      select(patid_cat, n, percent)

# table ----
knitr::kable(patid2, align = "c", caption = "Encrypted Patient ID")

## 3. Gender
```{r}
gender2 <- oregon_df %>%
  select(gender) %>%
  group_by(gender) %>%
  summarise(n = n()) %>%
  mutate(total_n = sum(n),
         percent = round(n/total_n, 4)*100,
         gender_cat = ifelse(gender == "F", "Female",
                      ifelse(gender == "M", "Male", 
                      ifelse(gender == "U", "Unknown", NA)))) %>%
    select(gender, gender_cat, n, percent)

knitr::kable(gender2, align = "c", caption = "Gender")

## Age
```{r}
oregon_df$yob <- as.numeric(oregon_df$yob)
age2 <- oregon_df %>%
  select(yob) %>%
  mutate(age = 2013-yob) %>%
  mutate(age_ind = ifelse(age < 15, 0,
                   ifelse(age >= 15 & age < 65, 1,
                   ifelse(age >= 65 & age <=120, 2, NA)))
  ) %>%
  group_by(age_ind) %>% 
  summarise(n = n()) %>% 
  mutate(total_n = sum(n),
         percent = round(n/total_n, 4)*100,
         age_cat = ifelse(age_ind == 0, "Age <15",
                   ifelse(age_ind == 1, "Age 15 to 65",
                   ifelse(age_ind == 2, "Age >= 65",
                   ifelse(is.na(age_ind), "Missing", NA))))) %>% 
  select(age_cat, age_ind, n, percent)

age2[4,1] <- age2[4,2] <- "Missing" # change NA to "Missing"

knitr::kable(age2, align = "c", caption = "Age")

## Race
```{r}
race2 <- oregon_df %>%
  select(race) %>%
  group_by(race) %>%
  summarise(n = n()) %>%
  mutate(total_n = sum(n),
         percent = round(n/total_n, 4)*100,
         race_cat = ifelse(race == 1, "White",
                    ifelse(race == 2, "Black or African American",
                    ifelse(race == 3, "American Indian or Alaska Native",
                    ifelse(race == 4, "Asian",
                    ifelse(race == 5, "Native Hawaiian or Pacific Islander",
                    ifelse(race == 6, "Some other race",
                    ifelse(race == 7, "Two or more races",
                    ifelse(race == 9, "Unknown",
                    ifelse(race == "", "Missing", NA)))))))))) %>%
    select(race, race_cat, n, percent)

knitr::kable(race2, align = "c", caption = "Race")

## Enthicity
```{r}
ethn2 <- oregon_df %>%
  select(ethn) %>%
  group_by(ethn) %>%
  summarise(n = n()) %>%
  mutate(total_n = sum(n),
         percent = round(n/total_n, 4)*100,
         ethn_cat = ifelse(ethn == 1, "Hispanic",
                    ifelse(ethn == 2, "Not Hispanic",
                    ifelse(ethn == 3, "Unknown",
                    ifelse(ethn == "", "Missing", NA))))) %>%
    select(ethn, ethn_cat, n, percent)

knitr::kable(ethn2, align = "c", caption = "Ethnicity")

## State
```{r}
state2 <- oregon_df %>%
  select(STATE) %>%
  group_by(STATE) %>%
  summarise(n = n()) %>%
  mutate(total_n = sum(n),
         percent = round(n/total_n, 4)*100,
         state_cat = ifelse(STATE == "OR", "Oregon",
                     ifelse(STATE != "OR"&is.na(STATE)==F, "Not Oregon",NA))) %>%
    select(STATE, state_cat, n, percent)

knitr::kable(state2, align = "c", caption = "State")

## ZIP
```{r}
zip2 <- oregon_df %>%
  select(ZIP) %>%
  group_by(ZIP) %>%
  summarise(n = n()) %>%
  mutate(total_n = sum(n),
         percent = round(n/total_n, 4)*100,
         zip_cat = ifelse(is.na(ZIP) == 0, "Non-Missing",
                   ifelse(is.na(ZIP) != 0, "Missing", NA))) %>%
    select(ZIP, zip_cat, n, percent)

knitr::kable(zip2, align = "c", caption = "ZIP COde")

## MSA
```{r}
msa2 <- oregon_df %>%
  select(MSA) %>%
  mutate(msa_cat = ifelse(MSA != "", "Non-Missing",
                   ifelse(MSA == "", "Missing", NA))) %>% 
  group_by(msa_cat) %>%
  summarise(n = n()) %>%
  mutate(total_n = sum(n),
         percent = round(n/total_n, 4)*100) %>%
  select(msa_cat, n, percent)

summary(as.factor(oregon_df$MSA))

knitr::kable(msa2, align = "c", caption = "MSA")

## Pos
```{r}
pos2 <- oregon_df %>%
  select(pos) %>%
  group_by(pos) %>%
  summarise(n = n()) %>%
  mutate(total_n = sum(n),
         percent = round(n/total_n, 4)*100,
         Place_of_service = ifelse(pos == "20", "Urgent Care Facility",
                            ifelse(pos == "23", "Emergency Room – Hospital", 
                            ifelse(is.na(pos) == 0, "Not Urgent", 
                            ifelse(is.na(pos) ==1, "Missing", NA))))) %>%
    select(pos, Place_of_service, n, percent)

knitr::kable(pos2, align = "c", caption = "Place of Service")

## from date
```{r}
fromdate2 <- oregon_df %>%
  select(fromdate) %>%
  mutate(dates = as.Date(fromdate2$fromdate, "%Y-%m-%d"))

summary(fromdate2)
hist(fromdate2$dates, "days", freq = TRUE)

## to date
```{r}
todate2 <- oregon_df %>%
  select(todate) %>%
  mutate(ddd = as.Date(todate2$todate, "%Y-%m-%d"))

todate3 <- oregon_df %>%
  filter (todate > "2014-12-31")
todate3

which(oregon_df$todate=="2299-12-31") # 129192

summary(todate2)

hist(todate2$ddd, "days", freq = TRUE)

## Length of stay
```{r}

## summary statistics key variables for study 
(date, dx, need to know urgent and emergency department visits, sex, age, patientid, state of residence, zipcode, county, race)
```{r}
summary(as.factor(oregon_df$personkey))
summary(as.factor(oregon_df$patid))
summary(as.factor(oregon_df$gender))
summary(as.factor(2013-oregon_df$yob)) # birth year -> age
summary(as.factor(oregon_df$race))
summary(as.factor(oregon_df$ethn))
summary(as.factor(oregon_df$STATE))
summary(as.factor(oregon_df$ZIP))
summary(as.factor(oregon_df$MSA)) 
summary(as.factor(oregon_df$fromdate))
summary(as.factor(oregon_df$todate))
summary(as.factor(oregon_df$pos)) # 23&20?
summary(as.factor(oregon_df$los))

## 1) *personkey*, Unique person identifier
person's ID, maximum is 14381640
length: 8
```{r personkey}
summary(as.factor(oregon_df$personkey))
min(oregon_df$personkey)
max(oregon_df$personkey) # seems length is 8
```
